
```{r core, message=F, echo=FALSE, warning=FALSE}
library(knitr)
require(knitcitations)
library(microbiome)
library(phyloseq)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(dada2)
theme_set(theme_bw(20))
knitr::opts_chunk$set(fig.width=10, fig.height=10, message=FALSE, warning=FALSE)
core_detection <- .1/100
core_prevalence <- 50/100
# Was created with: source("create_phyloseq.R")
phy20.1 <- readRDS("data/processed/phyloseq/phy20.1.RDS")
```



# Core microbiota.

Core microbiota is here defined based on the following parameters:

  * Detection threshold (relative abundance): `r 100 * core_detection`%
  * Prevalence threshold (above threshold in the population) `r 100 * core_prevalence`%.

The following taxonomic groups are in the core genera. Mean relative abundance and population prevalence (above detection threshold) are shown.

```{r core_genera, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, out.width="50%"}
ps <- aggregate_taxa(phy20.1, "genus")
ps <- microbiome::transform(ps, "compositional")
ps.core <- core(ps, detection = core_detection, prevalence = core_prevalence)
m <- round(100 * rowMeans(abundances(ps.core)), 1)
median <- round(100 * apply(abundances(ps.core), 1, median), 1)
prev <- round(100 * prevalence(ps.core, detection = core_detection), 1)
d <- as.data.frame(list(Taxon = names(m), Abundance_mean = m, Abundance_median = median, Prevalence = prev))
rownames(d) <- NULL
d <- d %>% arrange(desc(Abundance_mean), desc(Prevalence))
colnames(d) <- c("Taxon", "Mean Relative abundance (%)", "Median Relative abundance (%)", "Prevalence (%)")
kable(d)
```


Abundance variation across samples for each core taxa.

```{r core_rank_abundance, echo=FALSE, message=FALSE, fig.width=6, fig.height=4, fig.show="hold"}
theme_set(theme_bw(base_size = 13))
dfm <- melt(abundances(ps.core))
dfm$Var1 <- factor(dfm$Var1, levels = rev(d$Taxon))
dfm$Var_new = str_wrap(dfm$Var1, width = 40)
p2 <- ggplot(dfm, aes(x = Var_new, y = value)) +
       geom_boxplot() +
       geom_jitter(width = 0.1) +       
       scale_y_log10(label = scales::percent) +
       coord_flip() + theme(legend.position="none") +
       labs(y = "Relative abundance (%)", x = '', title = "Core abundance variation")
print(p2)
```


Core with varying detection and prevalence thresholds.

```{r core_heatmap, echo=FALSE, message=FALSE}
theme_set(theme_bw(base_size = 11))
library(RColorBrewer)
prevalences <- seq(.05, 1, .05)
detections <- 10^seq(log10(1e-3), log10(.2), length = 10)
ps <- microbiome::transform(ps, "compositional")
p1 <- plot_core(ps, plot.type = "heatmap", 
             prevalences = prevalences,
             detections = detections,
         colours = rev(brewer.pal(11, "Spectral")),
         min.prevalence = .2, horizontal = FALSE)
print(p1)
```
# Phylum abundance distributions

Abundance-prevalence plot per Phylum.
```{r phyla, echo=FALSE, message=FALSE, fig.width=12, fig.height=6, fig.show="keep", out.width="80%"}
library(reshape2)
a <- aggregate_taxa(phy20.1, level = "phylum")
xc <- abundances(transform(a, "compositional"))
dfm <- melt(xc)
names(dfm) <- c("phylum", "sample", "value")
dfm$Phylum <- factor(dfm$phylum,
    levels = as.character(unlist(dfm %>% group_by(phylum) %>%
  summarise(mean = mean(value)) %>%
  arrange(desc(mean)) %>%
  select(phylum))))
# Add small constant on zero values
dfm$value[dfm$value == 0] <- min(dfm$value[dfm$value > 0])/2
# Plot abundance distribution for the phyla
p <- ggplot(dfm, aes(x = value, fill = phylum)) +
         geom_density(alpha = 0.5) +
	 scale_x_log10(label = scales::percent) +
	 labs(x = "Relative abundance (%)", y = "Frequency",
	      title = "Phylum abundance distribution")
print(p)
```

Phylum abundance table (relative abundance %):

```{r core_phyla, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, out.width="50%"}
ps <- aggregate_taxa(phy20.1, "phylum")
ps <- microbiome::transform(ps, "compositional")
ps.core <- core(ps, detection = core_detection, prevalence = core_prevalence)
m <- round(100 * rowMeans(abundances(ps.core)), 1)
median <- round(100 * apply(abundances(ps.core), 1, median), 1)
prev <- round(100 * prevalence(ps.core, detection = core_detection), 1)
d <- as.data.frame(list(Taxon = names(m), Abundance_mean = m, Abundance_median = median, Prevalence = prev))
rownames(d) <- NULL
d <- d %>% arrange(desc(Abundance_mean), desc(Prevalence))
colnames(d) <- c("Taxon", "Mean Relative abundance (%)", "Median Relative abundance (%)", "Prevalence (%)")
kable(d)
```

```{r phylum_table, echo=FALSE, message=FALSE, fig.width=12, fig.height=6, fig.show="keep", out.width="100%"}
library(knitr)
a <- aggregate_taxa(phy20.1, level = "genus")
xc <- abundances(transform(a, "compositional"))
df <- melt(xc)
names(df) <- c("Genus", "sample", "value")
dfm <- df %>% group_by(Genus) %>% 
         summarise(mean = round(100 * mean(value), 1),
	           median = round(100 * median(value), 1),
		   min = round(100 * min(value), 1),
		   max = round(100 * max(value), 1)) %>%
	 arrange(desc(mean))
kable(dfm)
write.csv(dfm, file="genus table.csv")
```
