
```{r core, message=F, echo=FALSE, warning=FALSE}
library(knitr)
require(knitcitations)
library(microbiome)
library(phyloseq)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(dada2)
theme_set(theme_bw(20))
knitr::opts_chunk$set(fig.width=10, fig.height=10, message=FALSE, warning=FALSE)
core_detection <- .1/100
core_prevalence <- 50/100
# Was created with: source("create_phyloseq.R")
phy20.1 <- readRDS("data/processed/phyloseq/phy20.1.RDS")
```



# Core microbiota.

Core microbiota is here defined based on the following parameters:

  * Detection threshold (relative abundance): `r 100 * core_detection`%
  * Prevalence threshold (above threshold in the population) `r 100 * core_prevalence`%.

The following taxonomic groups are in the core genera. Mean relative abundance and population prevalence (above detection threshold) are shown.

```{r core_genera, echo=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=8, out.width="50%"}
ps <- aggregate_taxa(phy20.1, "genus")
ps <- microbiome::transform(ps, "compositional")
ps.core <- core(ps, detection = core_detection, prevalence = core_prevalence)
m <- round(100 * rowMeans(abundances(ps.core)), 1)
median <- round(100 * apply(abundances(ps.core), 1, median), 1)
prev <- round(100 * prevalence(ps.core, detection = core_detection), 1)
d <- as.data.frame(list(Taxon = names(m), Abundance_mean = m, Abundance_median = median, Prevalence = prev))
rownames(d) <- NULL
d <- d %>% arrange(desc(Abundance_mean), desc(Prevalence))
colnames(d) <- c("Taxon", "Mean Relative abundance (%)", "Median Relative abundance (%)", "Prevalence (%)")
kable(d)
```


Abundance variation across samples for each core taxa.

```{r core_rank_abundance, echo=FALSE, message=FALSE, fig.width=6, fig.height=4, fig.show="hold"}
theme_set(theme_bw(base_size = 13))
dfm <- melt(abundances(ps.core))
dfm$Var1 <- factor(dfm$Var1, levels = rev(d$Taxon))
dfm$Var_new = str_wrap(dfm$Var1, width = 40)
p2 <- ggplot(dfm, aes(x = Var_new, y = value)) +
       geom_boxplot() +
       geom_jitter(width = 0.1) +       
       scale_y_log10(label = scales::percent) +
       coord_flip() + theme(legend.position="none") +
       labs(y = "Relative abundance (%)", x = '', title = "Core abundance variation")
print(p2)
```


Phylum abundance table (relative abundance %):

```{r core_phyla, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, out.width="50%"}
ps <- aggregate_taxa(phy20.1, "phylum")
ps <- microbiome::transform(ps, "compositional")
ps.core <- core(ps, detection = core_detection, prevalence = core_prevalence)
m <- round(100 * rowMeans(abundances(ps.core)), 1)
#median <- round(100 * apply(abundances(ps.core), 1, median), 1)
prev <- round(100 * prevalence(ps.core, detection = core_detection), 1)
d <- as.data.frame(list(Taxon = names(m), Abundance_mean = m, Prevalence = prev))
rownames(d) <- NULL
d <- d %>% arrange(desc(Abundance_mean), desc(Prevalence))
colnames(d) <- c("Taxon", "Mean Relative abundance (%)", "Prevalence (%)")
kable(d)
```

```{r phylum_table, echo=FALSE, message=FALSE,  warning=FALSE, fig.width=12, fig.height=6, fig.show="keep", out.width="50%"}
library(knitr)
a <- aggregate_taxa(phy20.1, level = "genus")
xc <- abundances(transform(a, "compositional"))
df <- melt(xc)
names(df) <- c("Genus", "sample", "value")
dfm <- df %>% group_by(Genus) %>% 
         summarise(mean = round(100 * mean(value), 1),
	           #median = round(100 * median(value), 1),
		   #min = round(100 * min(value), 1),
		   #max = round(100 * max(value), 1),
		   prevalence = round(100 * prevalence(value), 1))%>%
	 arrange(desc(mean))
#kable(dfm) 
#write.csv(dfm, file="genus table.csv")
library(data.table)
a <- data.table:::print.data.table(dfm, nrows = 18, topn=18)
```

```{r core_abundance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=7, fig.show="hold"}
theme_set(theme_bw(base_size = 18))
core_detection1 <- .1/100
core_prevalence <- 20/100
b <- aggregate_taxa(phy20.1, "genus")
b <- microbiome::transform(b, "compositional")
b.core <- core(b, detection = core_detection1, prevalence = core_prevalence)
bs <- melt(abundances(b.core))
bs$Var2 <- factor(bs$Var2, levels = rev(d$Taxon))
bs$Var3 = str_wrap(bs$Var2, width =30)
p4 <- ggplot(bs, aes(x = Var1, y = value)) +
       geom_boxplot() +
       geom_jitter(width = 0.1) +       
       scale_y_log10(label = scales::percent) +
       coord_flip() + theme(legend.position="none") +
       labs(y = "Relative abundance (%)", x = '', title = "Core abundance variation")
print(p4)
```



```{r core_prevalance, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7, fig.show="hold"}
theme_set(theme_bw(base_size = 20))
#sorted.names <- c("Acinetobacter", "Aeromonas", "Anaerococcus", "Arthrobacter", "Bacillus", "Brevibacillus", "Brevibacterium", "Clostridium", "Corynebacterium", "Enterobacter", "Exiguobacterium", "Geobacillus", "Intestinimonas", "Micrococcus", "Moraxella", "Oceanobacillus", "Paenibacillus", "Pantoea", "Propionibacterium", "Pseudomonas", "Salmonella" ,"Staphylococcus", "Thiolamprovum", "Virgibacillus")

x <- prevalence(b.core)
df <- data.frame(prevalence = x) %>% arrange(prevalence) # Also sort by prevalence
df$Genus <- rownames(df)
sorted.names <- rownames(df) # Use the same order than in prevalences that were now sorted by magnitude
# sorted.names <- sort(rownames(df)) # Alphabetic order

df$Genus <- factor(df$Genus, levels=sorted.names)
ggplot(df, aes(x = df$Genus, y = prevalence )) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1)) +
  labs(y = "Prevalence (%)", x = 'Genus', title = "Core prevalence variation") 
```