---
title: "CSTAnalysis_SkinSamples.Rmd"
author: "Renuka"
date: "`r Sys.Date()`"
output: html_document
---


```{r initCST_CLR, message=F, echo=F, echo=FALSE}
library("phyloseq")
library("ggplot2")
library("cluster")
library("igraph")
library("markovchain")
library("RColorBrewer")
library(microbiome)
library("gridExtra")
library(tidyr)
library(dplyr)
set.seed(455423)
default.par <- par(no.readonly = TRUE)
options(stringsAsFactors = FALSE)
theme_set(theme_bw())
# EXCLUDEMARGINAL is a flag to exclude marginally preterm births (=37 weeks) in the later analysis
#EXCLUDEMARGINAL = TRUE
```
# Transform the data (proportions)
```{r transform-data_CLRleo, message=FALSE, echo=FALSE, warning=FALSE}
#site <- "Vaginal_Swab"
ps <- readRDS("data/processed/phyloseq/phy20.1.RDS")
tt <- data.frame(tax_table(ps))
ps <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps1<- prune_taxa(names(sort(taxa_sums(ps), T))[1:50], ps)
taxa.order1<- taxa_names(ps1)
```

# Cluster into CSTs
```{r MDS_CLRleo, warning=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
braydist <- phyloseq::distance(ps, method="bray")
ord = ordinate(ps, method = "MDS", distance = braydist)
plot_scree(ord) + xlim(as.character(seq(1,15))) + ggtitle("MDS-bray ordination eigenvalues")
evs <- ord$value$Eigenvalues
print(evs[1:20])
print(tail(evs))
```

## Denoise distance matrix
```{r PCoA-cutoff2_CLRleo, warning=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
h_sub5 <- hist(evs[6:length(evs)], 100)
plot(h_sub5$mids, h_sub5$count, log="y", type='h', lwd=10, lend=2)
```

## Determine number of clusters
We will use the gap statistic to indicate the number of clusters in this data:
```{r gap-stat_CLRleo, message=FALSE, echo=FALSE, warning=FALSE}
NDIM <- 7
x <- ord$vectors[,1:NDIM]  # rows=sample, cols=MDS axes, entries = value
pamPCoA = function(x, k) {
  list(cluster = pam(x[,1:2], k, cluster.only = TRUE))
}
gs = clusGap(x, FUN = pamPCoA, K.max = 12, B = 50)
plot_clusgap(gs) + scale_x_continuous(breaks=c(seq(0, 12, 2)))
```

## Cluster into CSTs

Perform PAM 3-fold clusters:

```{r pam-k5_CLRleo, message=FALSE, echo=FALSE, warning=FALSE}
K <- 3
x <- ord$vectors[,1:NDIM]
clust <- as.factor(pam(x, k=K, cluster.only=T))

# SWAPPING THE ASSIGNMENT OF 2 AND 3 TO MATCH RAVEL CST ENUMERATION
clust[clust==2] <- NA
clust[clust==3] <- 2
clust[is.na(clust)] <- 3
sample_data(ps)$CST <- clust
CSTs <- as.character(seq(K))
```

## Evaluate clustering
```{r see-pam-k3_CLRleo,eval=T,results = 'hide',  message=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
CSTColors <- brewer.pal(6,"Paired")[c(1,3,2,5,4,6)] # Length 6 for consistency with pre-revision CST+ coloration
names(CSTColors) <- CSTs
CSTColorScale <- scale_colour_manual(name = "CST", values = CSTColors[1:5])
CSTFillScale <- scale_fill_manual(name = "CST", values = CSTColors[1:5])
#grid.arrange(phyloseq::plot_ordination(ps, ord, color="CST") + CSTColorScale,plot_ordination(ps, ord, axes=c(3,4), color="CST") + CSTColorScale, main="Ordination by Cluster")
ok<-ordinate(ps, method="NMDS", distance=braydist)
knitr::opts_chunk$set(results = "hide")
```
```{r see-pam-k5u_CLRleo, message=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
plot_ordination(ps,ok , color="CST") + CSTColorScale + ggtitle("NMDS -- bray -- By Cluster")
```

#### Ordination plot for the Diet group

```{r see-pam-k5Sampesl55_CLRleo11, message=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
theme_set(theme_bw(base_size = 15))
o<- ordinate(ps, method = 'NMDS', distance = braydist)
ps1<- prune_samples(sample_data(ps)$Diet == "Veg", ps)
#o1<- ordinate(ps1, method="NMDS", distance=braydist)
m1<- meta(ps1)
oo1<- o$points[rownames(m1),]
oo1<- as.data.frame(oo1)
ps2<- prune_samples(sample_data(ps)$Diet == "Mixed", ps)
m2<- meta(ps2)
#o2<- ordinate(ps2, method="NMDS", distance=braydist)
oo2<- o$points[rownames(m2),]
oo2<- as.data.frame(oo2)
knitr::opts_chunk$set(results = "hide")
```

```{r CLRleo2233, fig.height = 9, fig.width = 15, , message=FALSE, echo=FALSE, warning=FALSE}
theme_set(theme_bw(base_size = 10))
p1<- list()
 m1<-meta(ps)$Veg
for(i in levels(m1)) {
  pshm1 <- prune_samples(sample_data(ps1)$Diet == "Veg", ps1)
  n<- rownames(meta(pshm1))
  df1<-oo1[n,]
  df<- cbind(df1, meta(pshm1))
  p1[[i]]<-ggplot(df, aes(x = MDS1, y = MDS2, color = CST))+ geom_point(size= 1.5)+ labs(title = paste0(i), subtitle = '') +  xlim(-1, 1)+ ylim(-.5, .5)
  #print(p1[[i]])
}
p2<- list()
for(i in levels(m1)) {
  pshm2 <- prune_samples(sample_data(ps2)$Diet == "Mixed", ps2)
  n<- rownames(meta(pshm2))
  df3<-oo2[n,]
  df2<-cbind(df3, meta(pshm2))
  p2[[i]]<-ggplot(df2, aes(x = MDS1, y = MDS2, color = CST))+ geom_point(size= 1.5)+ labs(title = paste0(i), subtitle = '') + xlim(-1, 1)+ ylim(-.5, .5)
  #print(p2[[i]])
}
```

```{r CLRleo_panel, fig.height=5, fig.width=15, message=FALSE, echo=FALSE, warning=FALSE, error=F}
theme_set(theme_bw(base_size = 10))
library(cowplot)
library(multipanelfigure)
#figure1 <- multi_panel_figure(columns = 1, rows = 2, panel_label_type = "none")
#figure1 %<>%
  #fill_panel(p1$"Veg", column = 1, row = 1) %<>%
  #fill_panel(p1$`2nd group`, column = 2, row = 1) %<>%
  #fill_panel(p1$`3rd group`, column = 3, row = 1) %<>%
  #fill_panel(p1$`4th group`, column = 2, row = 1) %<>%
  #fill_panel(p2$"Mixed", column = 1, row = 2) %<>%
  #fill_panel(p2$`2nd group`, column = 2, row = 2) %<>%
  #fill_panel(p2$`3rd group`, column = 3, row = 2) %<>%
  #fill_panel(p2$`4th group`, column = 2, row = 2)
#plot(figure1)
#par(mfrow=c(1,3), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
#ppp<-grid.arrange(p1$`1st group`, p1$`2nd group`, p2$`1st group`, p2$`2nd group`)
#pppp<-grid.arrange(p1$`3rd group`, p1$`4th group`, p2$`3rd group`, p2$`4th group`)
#p<- grid.arrange(ppp, pppp, ncol=2)
```

```{r see-pam-k3Sampesl_CLRleo, message=FALSE,, results="hide",  message=FALSE, echo=FALSE, warning=FALSE}
o<- ordinate(ps, method="NMDS", distance=braydist)
knitr::opts_chunk$set(results = "hide")
```

```{r see-pam-k3Sampesl55_CLRleo, message=FALSE, message=FALSE, echo=FALSE, warning=FALSE,fig.height=8, fig.width=15, out.width='50%'}
theme_set(theme_bw(base_size = 20))
plot_ordination(ps,o , "samples", color="Diet", shape = 'Veg') + geom_point(size=2)+ xlim(-1, 1)+ ylim(-.5, .5)
#plot_ordination(ps,o , "samples", color="InterventionGroup", shape = 'StVisit') + geom_point(size=2)+ xlim(-1, 1)+ ylim(-.5, .5)
par(mfrow=c(1,4), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
```

# Principal Coordinates Analysis (PCoA)

Coloured by DMM community type

```{r pcoa, echo=FALSE, message=FALSE, fig.width=9, fig.height=5, out.width="33%", fig.show="hold", warning=FALSE}
#theme_set(theme_bw(15))
method <- "PCoA"
trans <- "compositional"
distance <- "bray"
p <- plot_landscape(microbiome::transform(ps , "compositional"),
                    distance = distance,
                    method = method,
                    size = 3,
                    col = "CST",
                    shading = FALSE) +
  scale_colour_brewer(palette = "Set1")		    
print(p)
```

### Heatmap

Heatmaps for the community state types.

```{r clust-diverse_CLRleo, eval= T, message=FALSE, echo=FALSE, warning=FALSE, out.width="33%", fig.show="hold", warning=FALSE}
theme_set(theme_bw(base_size = 20))
library(BBmisc)
library(reshape2)
library("RColorBrewer")
library(tidyr)
set.seed(1234)

# Identify ASVs that are potentially different between CSTs
pvs <- c()
for (tax in taxa(ps)) {
  pvs[[tax]] <- kruskal.test(abundances(ps)[tax,] ~ meta(ps)$CST)$p.value
}
pvs <- sort(unlist(pvs))
significant.taxa <- names(pvs)[which(as.numeric(p.adjust(pvs)) < 0.25)]

# Pick full names for the taxa
full.names <- apply(tax_table(ps), 1, function (x) {paste(x, collapse = "_")})

pso<- prune_taxa(significant.taxa, ps)
pso1<- microbiome::transform(pso, 'clr')
g<- otu_table(pso1)
g1<-BBmisc::normalize(g@.Data, 'standardize', range = c(0,1))
colnames(g1)<- colnames(otu_table(pso1)@.Data)
j<- otu_table(g1, taxa_are_rows = T)
otu_table(pso1)<- j
p <- list()
sort.taxa <- neatsort(pso1, "rows")
for(CST in CSTs) {
  pshm <- prune_samples(sample_data(pso1)$CST == CST, pso1)
  dfm <- melt(abundances(pshm))
  colnames(dfm) <- c("Taxa", "Sample", "value")
  dfm$Taxa <- factor(dfm$Taxa, levels = sort.taxa)
  dfm$Sample <- factor(dfm$Sample, levels = neatsort(pshm, "cols"))
  p[[CST]]<- heat((dfm), "Sample","Taxa", "value", legend.text = "Abundance", order.rows = F, 
                order.cols = F, limits= c(-5,5), step = 0.5) +
		ggtitle(paste("CST:", CST))
print(p[[CST]])
}
```


### Boxplots for each genus in the three clusters

```{r clust-diverse_boxplot_CLRleo, message=FALSE, echo=FALSE, warning=FALSE, out.width="25%"}
theme_set(theme_bw(base_size = 20))
library(ggbeeswarm)
df<- meta(pso)

for (tax in rev(significant.taxa)) {

  # Pick the abundance values for the given taxonomic group ("tax")
  # Let us use CLR transformed values
  df$signal <- abundances(transform(pso, "clr"))[tax,]

  # Retrieve the full name for this taxon
  tax.full.name <- full.names[tax]

  # Plot taxon abundance (signal) against the cluster (CST) as a boxplot
  p <- ggplot(df, aes(x = CST, y = signal)) +
         labs(title = tax) +
         geom_boxplot() +
	 geom_beeswarm()
  print(p)
  
}
```


Table of full names for the taxa:

```{r full_names, message=FALSE, echo=FALSE, warning=FALSE}
tab <- data.frame(ASV = significant.taxa, Full_name = full.names[significant.taxa])
rownames(tab) <- NULL
kable(tab)
```


##### -----------------------------------------------------------------

What is the purpose of this analysis? Some comments would help.

```{r clust-diverse_table_CLR,eval=T, results='show', message=FALSE, echo=FALSE, warning=FALSE}
gg<- meta(ps)
df11<- subset(gg, CST==1 & Diet =="Veg")
df12<- subset(gg, CST==1& Diet =="Mixed")
df1<- subset(gg, CST==1); df1<- df1[, -seq(1:46)]
df2<- subset(gg, CST==2); df2<- df2[, -seq(1:46)]
df3<- subset(gg, CST==3); df3<- df3[, -seq(1:46)]
ddddf1<- c(round(dim(df11)[1]/dim(df1)[1] *100,1))
ddddf2<-c(round(dim(df12)[1]/dim(df1)[1]*100 ,1))
          
dfdf<- cbind(ddddf1, ddddf2)
rownames(dfdf)<- c('Clust.1')
colnames(dfdf)<- c('Veg','Mixed')
library(knitr)
kable(dfdf)
```

#### Fisher-exact test at the first and second visits between clusters

Do we have first and second visits?

```{r clust-diverse_table_CLR09,eval=T, results='show', message=FALSE, echo=FALSE, warning=FALSE}
gh<- xtabs(~  CST +Diet, data = gg)
addmargins(gh)
ff<- fisher.test(gh)
```

```{r clust-diverse_table_CLR0977,eval=T, results='show', message=FALSE, echo=FALSE, warning=FALSE}
print('Fisher exact test p-value is: ')
print(ff$p.value)
```
