
# Analysis

```{r alpha, message=F, echo=F, echo=FALSE, warning=FALSE }
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(phyloseq)
library(dada2)
library(readxl)
library(dplyr)
library(knitr)
library(knitcitations)
library(microbiome)
library(phyloseq)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(dada2)
library(dplyr)
theme_set(theme_bw(20))
knitr::opts_chunk$set(fig.width=10, fig.height=10, message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.path="figure_diet/")

# Was created with: source("create_phyloseq.R")
phy <- readRDS("data/processed/phyloseq/phy20.1.RDS")
```

# Alpha diversity analysis

```{r alpha1, echo=FALSE, message=FALSE, warning=FALSE}
library(microbiome)
opts_chunk$set(dev="CairoPNG")

# Estimate alpha diversities
A <- alpha(phy)
index <- "diversity_shannon"
```


# Group-wise comparisons
* Diversity index: `r index`

```{r group_comp, echo=FALSE, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, fig.show="keep", out.width="50%"}
library(Cairo)
df <- meta(phy)
df <- bind_cols(df, A)
df$index <- df[[index]]
pv <- wilcox.test(data = df, index ~ factor(Diet))$p.value
library(ggbeeswarm)
p1 <- ggplot(df, aes(x = Diet, y = index, fill = Diet)) +
        ggbeeswarm::geom_beeswarm() +
	theme(axis.text.x = element_text(angle = 0, hjust = 1)) +
        labs(y = "Shannon Diversity",
	     x = "",
	     title = paste0("Wilcoxon test p=", round(pv, 2)))

print(p1)
```

# Ordination

```{r init, echo=FALSE, message=FALSE}
library(microbiome)
library(phyloseq)
library(vegan)
opts_chunk$set(dev="CairoPNG")
method <- "PCoA"
trans <- "compositional"
distance <- "jaccard"
```

## *Principal Coordinates Analysis (PCoA)*

* Ordination method: `r method`
* Dissimilarity measure: `r distance`

```{r pcoa, echo=FALSE, message=FALSE, fig.width=9, fig.height=5, out.width="50%", fig.show="hold"}
theme_set(theme_bw(15))
library(microbiome)
p <- plot_landscape(microbiome::transform(phy , "compositional"),
                    distance = "jaccard",
                    method = "PCoA",
                    size = 3,
                    col = "Diet",
                    shading = FALSE) +
  scale_colour_brewer(palette = "Set1")		    
print(p)
```

## Differential abundance analysis 

Differential abundance analysis is carriedout using DESeq2 method

```{r DESeq2, echo=FALSE, message=FALSE, fig.width=9, fig.height=5, out.width="50%", fig.show="hold"}
library(phyloseq)
library(reshape2)
library(DESeq2)
padj.th <- 0.25
mypseq <- phy

# Remove NAs
group <- "Diet"
keep <- meta(phy)[, group] %>% complete.cases()

df <- meta(mypseq)[keep, ]

mygroup <- "Diet"
df$group <- as.factor(df[[group]] == mygroup)
sample_data(mypseq) <- sample_data(df)
quiet(ds2 <- phyloseq_to_deseq2(mypseq, ~ Diet))
dds <- DESeq(ds2)
res <- results(dds)
df <- as.data.frame(res)
df$taxon <- rownames(df)
#df <- df %>% arrange(desc(abs(log2FoldChange)), padj) %>%
#             filter(padj < padj.th) 
df <- df %>% arrange(padj, desc(abs(log2FoldChange))) %>%
             filter(padj < padj.th) 
df <- subset(df, select= c(log2FoldChange, padj, taxon))
rownames(df) <- NULL

# WHat is the purpose of this one?
#full.names <- apply(tax_table(phy), 1, function (x) {paste(x, collapse = "_")})
# top.taxa <- as.character(df$taxon)
# Add full phylogenetic name for this OTU
df$full_name <- full.names[df$taxon]
library(knitr)
kable(df)


df <- meta(phy)
ps <- microbiome::transform(phy, "clr")
for (tax in top.taxa) {
  df$taxa <- abundances(ps)[tax, ]
  p <- ggplot(df, aes(x = Diet, y = taxa)) +
    geom_boxplot() + 
    ggbeeswarm::geom_beeswarm() +
    theme(axis.text.x = element_text(angle = 360, hjust = 1)) +
    labs(x = "Diet", y = "Abundance", title = tax)
  print(p)
}
```

# PERMANOVA analysis

p value for the effect of diet is (p=0.31), which is not significant.

```{r PERMANOVA, echo= FALSE}
library(microbiome)
library(ggplot2)
library(dplyr)

pseq <- phy# Rename the example data
data = microbiome::meta(phy)
# Pick relative abundances (compositional) and sample metadata 
pseq.rel <- microbiome::transform(pseq, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)
library(vegan)
library(phyloseq)
permanova <- adonis(t(otu) ~ Diet,
               data = meta(phy), permutations=99, method = "bray")

# P-value
print(as.data.frame(permanova$aov.tab)["Diet", "Pr(>F)"])

dist <- vegdist(t(otu))
anova(betadisper(dist, meta$Diet))
```

# Investigate the top factors

```{r top_factors, echo=FALSE, fig.width=20, fig.height=10}
library(vegan)
coef1 <- coefficients(permanova)["Diet1", ]
top.coef1 <- coef1[rev(order(abs(coef1)))[1:20]]
names(top.coef1) <- full.names[names(top.coef1)]
par(mar = c(3, 14, 2, 1))
barplot(sort(top.coef1), horiz = T, las = 1, main = "Top taxa / 1")
```