
# Analysis

```{r alpha, message=F, echo=F, echo=FALSE, warning=FALSE}
library(knitr)
library(knitcitations) 
library(microbiome)
library(phyloseq)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(dada2)
library(dplyr)
library(microbiome)
library(microbiome)
library(phyloseq)
library(vegan)
theme_set(theme_bw(20))
knitr::opts_chunk$set(fig.width=10, fig.height=10, message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.path="figure_location/")
opts_chunk$set(dev="CairoPNG")
# Was created with: source("create_phyloseq.R")
phy <- readRDS("data/processed/phyloseq/phy20.1.RDS")
method <- "PCoA"
trans <- "compositional"
distance <- "bray"
index <- "diversity_shannon"
```


## Alpha diversity analysis

Diversity index: `r index`

```{r alpha1, echo=FALSE, message=FALSE, warning=FALSE}
# Estimate alpha diversities
A <- alpha(phy)
b <- meta(phy)
b2 <- bind_cols(b,A)
is.data.frame(b) 
is.data.frame(b2)
sapply(split(b2$diversity_shannon, b2$Geographical_location), mean)
```


```{r group_comp, echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6, fig.show="keep", out.width="50%"}
theme_set(theme_bw(base_size=25))
library(Cairo)
df <- meta(phy)
df <- bind_cols(df, A)
df$index <- df[[index]]
pv <- kruskal.test(data = df, index ~ factor(Geographical_location))$p.value
library(ggbeeswarm)
p1 <- ggplot(df, aes(x = Geographical_location, y = index)) +
       geom_boxplot() +
       geom_jitter(width = 0.1) +  
        theme(axis.text.x = element_text(angle = 0, hjust = 1)) + theme(legend.position = "none")+
        labs(y = "Shannon Diversity",
	     x = "",
	     title = "Location")
print(p1)
```  



## Differential abundance analysis (with ANCOM)

In [jointanalysis.md](jointanalysis.md) it was shown that geographical location has a significant effect.

Here, we investigate individual taxonomic groups in more detail.

For community comparison, see [CSTAnalysis_SkinSamples.md](CSTAnalysis_SkinSamples.md)

Significant (or marginally significant) taxa between geographical locations.

```{r diffab, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=5, out.width="25%", fig.show="hold"}
theme_set(theme_bw(base_size=25))
library(readr)
library(tidyverse)
otu_data <- abundances(phy)
meta_data <- meta(phy)
# To get this, clone
# https://github.com/FrederickHuangLin/ANCOM.git
source("ANCOM/scripts/ancom_v2.1.R")
# Step 1: Data preprocessing
feature_table = otu_data;
sample_var = "Sample";
group_var = NULL
out_cut = 0.05;
zero_cut = 0.90;
lib_cut = 1000;
neg_lb = FALSE
prepro = feature_table_pre_process(feature_table,
                                   meta_data,
				   sample_var,
				   group_var, 
                                   out_cut,
				   zero_cut,
				   lib_cut,
				   neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
# Step 2: ANCOM
main_var <- "Geographical_location";
p_adj_method <- "BH";
alpha <- 0.05
adj_formula <- NULL;
rand_formula <- NULL
res <- ANCOM(feature_table, meta_data, struc_zero,
             main_var, p_adj_method, 
             alpha, adj_formula, rand_formula)
res.sorted <- res$out %>% arrange(desc(W))
top.taxa <- res.sorted[which(res.sorted[,3]), "taxa_id"]
head(res.sorted)
ps <- microbiome::transform(phy, "compositional")
d <- meta(ps)
for (tax in top.taxa) {
  d$taxa <- abundances(ps)[tax, ]
  # d$gen <- taxa(ps)["genus", ]
  
  p <- ggplot(d, aes(x = Geographical_location, y = taxa)) +
    geom_boxplot() + 
    ggbeeswarm::geom_beeswarm() +
    scale_y_log10()+
    theme(axis.text.x = element_text(angle = 360, hjust = 1)) +
    #labs(x = "Location", y = "Abundance (CLR)", title = map_levels(tax, from='tax', to='genus',ps))
    labs(x = "Location", y = "Abundance (CLR)", title = as.vector(tax_table(ps)[tax, "genus"]))
  print(p)
}
```


```{r diffab, echo=FALSE, message=FALSE, warning=FALSE}
s1 <- sample_names(subset_samples(ps, Geographical_location == "Ahmednagar"))
s2 <- sample_names(subset_samples(ps, Geographical_location == "Nashik"))
s3 <- sample_names(subset_samples(ps, Geographical_location == "Pune"))
rp<- rowMeans(abundances(pseq)[, s1])/rowMeans(abundances(pseq)[, s2])/rowMeans(abundances(pseq)[, s3])
sample_data(ps)$rp<- rowMeans(abundances(pseq)[, s1])/rowMeans(abundances(pseq)[, s2])/rowMeans(abundances(pseq)[, s3])
```

