
##### -----------------------------------------------------------------

What is the purpose of this analysis? Some comments would help.

```{r clust-diverse_table_CLR,eval=T, results='show', message=FALSE, echo=FALSE, warning=FALSE}
gg<- meta(ps)
df11<- subset(gg, CST==1 & Diet =="Veg")
df12<- subset(gg, CST==1& Diet =="Mixed")
df1<- subset(gg, CST==1); df1<- df1[, -seq(1:46)]
df2<- subset(gg, CST==2); df2<- df2[, -seq(1:46)]
df3<- subset(gg, CST==3); df3<- df3[, -seq(1:46)]
ddddf1<- c(round(dim(df11)[1]/dim(df1)[1] *100,1))
ddddf2<-c(round(dim(df12)[1]/dim(df1)[1]*100 ,1))
          
dfdf<- cbind(ddddf1, ddddf2)
rownames(dfdf)<- c('Clust.1')
colnames(dfdf)<- c('Veg','Mixed')
library(knitr)
kable(dfdf)
```

#### Fisher-exact test at the first and second visits between clusters

Do we have first and second visits?

```{r clust-diverse_table_CLR09,eval=T, results='show', message=FALSE, echo=FALSE, warning=FALSE}
gh<- xtabs(~  CST +Diet, data = gg)
addmargins(gh)
ff<- fisher.test(gh)
```

```{r clust-diverse_table_CLR0977,eval=T, results='show', message=FALSE, echo=FALSE, warning=FALSE}
print('Fisher exact test p-value is: ')
print(ff$p.value)
```


#### Ordination plot for the Diet group

```{r see-pam-k5Sampesl55_CLRleo11, message=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
theme_set(theme_bw(base_size = 15))
o<- ordinate(ps, method = 'NMDS', distance = braydist)
ps1<- prune_samples(sample_data(ps)$Diet == "Veg", ps)
m1<- meta(ps1)
oo1<- o$points[rownames(m1),]
oo1<- as.data.frame(oo1)
ps2<- prune_samples(sample_data(ps)$Diet == "Mixed", ps)
m2<- meta(ps2)
oo2<- o$points[rownames(m2),]
oo2<- as.data.frame(oo2)
knitr::opts_chunk$set(results = "hide")
```

```{r CLRleo2233, fig.height = 9, fig.width = 15, , message=FALSE, echo=FALSE, warning=FALSE}
theme_set(theme_bw(base_size = 10))
p1<- list()
m1<-meta(ps)$Veg

for(i in levels(m1)) {
  pshm1 <- prune_samples(sample_data(ps1)$Diet == "Veg", ps1)
  n<- rownames(meta(pshm1))
  df1<-oo1[n,]
  df<- cbind(df1, meta(pshm1))
  p1[[i]] <- ggplot(df, aes(x = MDS1, y = MDS2, color = CST)) +
    geom_point(size= 1.5) +
    labs(title = paste0(i), subtitle = '') +
    xlim(-1, 1) +
    ylim(-.5, .5)
}

p2<- list()
for(i in levels(m1)) {
  pshm2 <- prune_samples(sample_data(ps2)$Diet == "Mixed", ps2)
  n<- rownames(meta(pshm2))
  df3<-oo2[n,]
  df2<-cbind(df3, meta(pshm2))
  p2[[i]] <- ggplot(df2, aes(x = MDS1, y = MDS2, color = CST)) +
    geom_point(size= 1.5) +
    labs(title = paste0(i), subtitle = '') +
    xlim(-1, 1) +
    ylim(-.5, .5)
}
```

```{r CLRleo_panel, fig.height=5, fig.width=15, message=FALSE, echo=FALSE, warning=FALSE, error=F}
theme_set(theme_bw(base_size = 10))
library(cowplot)
library(multipanelfigure)
  #fill_panel(p2$`2nd group`, column = 2, row = 2) %<>%
  #fill_panel(p2$`3rd group`, column = 3, row = 2) %<>%
  #fill_panel(p2$`4th group`, column = 2, row = 2)
#plot(figure1)
#par(mfrow=c(1,3), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
#ppp<-grid.arrange(p1$`1st group`, p1$`2nd group`, p2$`1st group`, p2$`2nd group`)
#pppp<-grid.arrange(p1$`3rd group`, p1$`4th group`, p2$`3rd group`, p2$`4th group`)
#p<- grid.arrange(ppp, pppp, ncol=2)
```

```{r DESeq2, echo=FALSE, message=FALSE, fig.width=9, fig.height=5, out.width="33%", fig.show="hold"}
library(phyloseq)
library(reshape2)
library(DESeq2)
padj.th <- 0.25
mypseq <- phy

# Remove NAs
group <- "Geographical_location"
keep <- meta(phy)[, group] %>% complete.cases()

df <- meta(mypseq)[keep, ]

mygroup <- "Geographical_location"
df$group <- as.factor(df[[group]] == mygroup)
sample_data(mypseq) <- sample_data(df)
quiet(ds2 <- phyloseq_to_deseq2(mypseq, ~ Geographical_location))
dds <- DESeq(ds2)
res <- results(dds)
df <- as.data.frame(res)
df$taxon <- rownames(df)
#df <- df %>% arrange(desc(abs(log2FoldChange)), padj) %>%
#             filter(padj < padj.th) 

df <- df %>% arrange(padj, desc(abs(log2FoldChange))) %>%
             filter(padj < padj.th) 
df <- subset(df, select= c(log2FoldChange, padj, taxon))
rownames(df) <- NULL

# WHat is the purpose of this one?
full.names <- apply(tax_table(phy), 1, function (x) {paste(x, collapse = "_")})
# top.taxa <- as.character(df$taxon)

# Add full phylogenetic name for this OTU
df$full_name <- full.names[df$taxon]

library(knitr)
kable(df)


df <- meta(phy)
ps <- microbiome::transform(phy, "clr")
for (tax in top.taxa) {
  df$taxa <- abundances(ps)[tax, ]
  p <- ggplot(df, aes(x = Geographical_location, y = taxa)) +
    geom_boxplot() + 
    ggbeeswarm::geom_beeswarm() +
    theme(axis.text.x = element_text(angle = 360, hjust = 1)) +
    labs(x = "Geographical location", y = "Abundance", title = tax)
  print(p)
}
```

# Differential abundance analysis

```{r DESeq2, echo=FALSE, message=FALSE, fig.width=9, fig.height=5, out.width="25%", fig.show="hold"}
location <- meta(phy)$Geographical_location
A <- abundances(phy);
pv <- apply(A, 1, function (x) {kruskal.test(x ~ location)$p.value});
padj <- p.adjust(pv);

df <- data.frame(taxon = names(padj), padj = padj)
# Add full names for the taxa
full.names <- apply(tax_table(phy), 1, function (x) {paste(x, collapse = "_")})
df$full_name <- full.names[df$taxon]
df <- df %>% filter(padj < 0.25) %>% arrange(padj)
rownames(df) <- NULL

library(knitr)
kable(df)

ps <- microbiome::transform(phy, "clr")
d <- meta(ps)
for (tax in df$taxon) {
  d$taxa <- abundances(ps)[tax, ]
  p <- ggplot(d, aes(x = Geographical_location, y = taxa)) +
    geom_boxplot() + 
    ggbeeswarm::geom_beeswarm() +
    theme(axis.text.x = element_text(angle = 360, hjust = 1)) +
    labs(x = "Location", y = "Abundance (CLR)", title = tax)
  print(p)
}
```






# Community composition (with PERMANOVA)

```{r PERMANOVA}
library(microbiome)
library(ggplot2)
library(dplyr)

pseq <- phy# Rename the example data
data = microbiome::meta(phy)
# Pick relative abundances (compositional) and sample metadata 
pseq.rel <- microbiome::transform(pseq, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)
library(vegan)
library(phyloseq)
permanova <- adonis(t(otu) ~ Geographical_location,
               data = meta(phy), permutations=99, method = distance)

# P-value
print(as.data.frame(permanova$aov.tab)["Geographical_location", "Pr(>F)"])

## Check beta dispersion - is this OK?
#dist <- vegdist(t(otu))
#anova(betadisper(dist, meta$Geographical_location))
```
## Principal Coordinates Analysis (PCoA)

* Ordination method: `r method`
* Dissimilarity measure: `r distance`

```{r pcoa, echo=FALSE, message=FALSE, fig.width=9, fig.height=5, out.width="50%", fig.show="hold"}
theme_set(theme_bw(15))
library(microbiome)
p <- plot_landscape(microbiome::transform(phy , "compositional"),
                    distance = distance,
                    method = method,
                    size = 3,
                    col = "Geographical_location",
                    shading = FALSE) +
  scale_colour_brewer(palette = "Set1")		    
print(p)
```





# Investigate the top factors

```{r top_factors, fig.width=20, fig.height=10}
library(vegan)

coef1 <- coefficients(permanova)["Geographical_location1", ]
top.coef1 <- coef1[rev(order(abs(coef1)))[1:20]]
names(top.coef1) <- full.names[names(top.coef1)]

coef2 <- coefficients(permanova)["Geographical_location2", ]
top.coef2 <- coef2[rev(order(abs(coef2)))[1:20]]
names(top.coef2) <- full.names[names(top.coef2)]

par(mar = c(3, 20, 2, 1), mfrow = 2)
barplot(sort(top.coef1), horiz = T, las = 1, main = "Top taxa / 1")
barplot(sort(top.coef2), horiz = T, las = 1, main = "Top taxa / 2")
```

